---
title: "Spatial data wrangling"
author: "Julianna Renzi"
date: "2/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
require(sf)
require(tmap)
```

## California county outlines

Read it in as a shapefile (read_sf)

```{r}
ca_counties <- read_sf(here("data", "ca_counties", "CA_Counties_TIGER2016.shp"))

# wrangle (and note sticky geometry!)
ca_subset <- ca_counties %>% 
  select(NAME, ALAND) %>% 
  rename(county_name = NAME, land_area = ALAND)

# note: it just assumes you want to keep the spatial information, and you can work with the rest of the data as if it’s a non-spatial data frame (and the spatial information just “sticks” - hence the term “sticky geometry”). So even though we only called NAME and ALAND in the select() function, we see that the geometry column still exists!

# check and set the coordinate reference system
ca_subset %>% st_crs # WGS84

```
Look at it

```{r}
ggplot(data = ca_subset) +
  geom_sf(aes(fill = land_area), color = "white", size = 0.1) +
  theme_void() +
  scale_fill_gradientn(colors = c("cyan", "blue", "purple"))

```

## Invasive red sesbania records (spatial points)

```{r}
sesbania <- read_sf(here("data", "red_sesbania", "ds80.shp"))

# check the CRS:
sesbania %>% st_crs()
```

```{r}
# need to update the CRS to match
sesbania <- st_transform(sesbania, 3857)

# check it
sesbania %>% st_crs()
```

Plot them together

```{r}
ggplot() +
  geom_sf(data = ca_subset) +
  geom_sf(data = sesbania, size = 1, color = "red")
```

Let’s say we want to find the count of red sesbania observed locations in this dataset by county. How can I go about joining these data so that I can find counts? Don’t worry…st_join() has you covered for spatial joins!

```{r}
ca_sesbania <- ca_subset %>% 
  st_join(sesbania)
```

And then we can find counts (note: these are not counts for individual plants, but by record in the dataset) by county:

```{r}
sesbania_counts <- ca_sesbania %>% 
  count(county_name)

# plot a chloropleth using the number of records for red sesbania as the fill color (instead of what we used previously, land area)

ggplot(data = sesbania_counts) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) + 
  scale_fill_gradientn(colors = c("lightgray", "orange", "red")) +
  theme_minimal() +
  labs(fill = "Number of S. punicea records")
```

So we see that we can still use our usual wrangling skills! Let’s do a bit more for fun, just to prove that our existing wrangling skills still work with spatial data - the spatial information just sticks to it! Only plot the county with the greatest number of red sesbania records (Solano), and make a map of those locations (yeah there are many ways to do this):

```{r}
# subset of sebania point locations only in Solano County
solano_sesbania <- sesbania %>% 
  filter(COUNTY == "Solano")

# only keep Solano polygon
solano <- ca_subset %>% 
  filter(county_name == "Solano")

ggplot() +
  geom_sf(data = solano) +
  geom_sf(data = solano_sesbania)
```
## Interactive map with tmap

Let’s make one for our California counties (fill aesthetic by land area) with the red sesbania locations on top

```{r}
# set viewing mode to interactive
tmap_mode(mode = "view")

# make map with the polygon fill color updated by variable 'land_area', updating the color palette to "BuGn"), then add another shape layer for the sesbania records (added as dots)
tm_shape(ca_subset) +
  tm_fill("land_area", palette = "BuGn") +
  tm_shape(sesbania) +
  tm_dots()


```
Check out cool vignettes on other cool things you can do with `tmap` here: https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html
