---
title: "Time series"
author: "Julianna Renzi"
date: "2/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
require(tsibble)
require(feasts)
require(fable)
require(here)
```

Read in energy data and convert to a tsibble

```{r}
energy <- read_csv(here("data", "energy.csv")) 

energy_ts <- energy %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date) # get it into tsibble format

```
## Exploration

Timeseries plot

```{r}
ggplot(data = energy_ts, aes(x = date, y = res_total)) +
  geom_line() + 
  labs(y = "Residential energy consumption \n (Trillion BTU)")
```

Season plot

```{r}
energy_ts %>% 
  gg_season(y = res_total) + 
  theme_minimal() +
  labs(x = "month",
       y = "residential energy consumption (trillion BTU)")
```
Subseries plot

```{r}
energy_ts %>% 
  gg_subseries(res_total)
```

Takeaway is similar: there is clear seasonality (higher values in winter months), with an increasingly evident second peak in June/July/August. This reinforces our takeaways from the raw data and seasonplots.

## Decomposition (by STL)

See Rob Hyndman’s section on STL decomposition to learn how it compares to classical decomposition we did last week: “STL is a versatile and robust method for decomposing time series. STL is an acronym for “Seasonal and Trend decomposition using Loess”, while Loess is a method for estimating nonlinear relationships."

Notice that it allows seasonality to vary over time (a major difference from classical decomposition, and important here since we do see changes in seasonality).

```{r}
# find STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))

# view the components
components(dcmp)

# visualize the decomposed components
components(dcmp) %>% autoplot() +
  theme_minimal()
```
Autocorrelation function (ACF)

```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot()

# We use the ACF to explore autocorrelation (here, we would expect seasonality to be clear from the ACF):

# And yep, we see that observations separated by 12 months are the most highly correlated, reflecting strong seasonality we see in all of our other exploratory visualizations.

```
## Forecasting by Holt-Winters exponential smoothing

Note: here we use ETS, which technically uses different optimization than Holt-Winters exponential smoothing, but is otherwise the same (From Rob Hyndman: “The model is equivalent to the one you are fitting with HoltWinters(), although the parameter estimation in ETS() uses MLE.”)

```{r}
# create the model (note "M" standed for multiplicative)
energy_fit <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M"))
  )

# forecast using the model 10 years into the future:
energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

# plot just the forecasted values (with 90 & 90% CI's)
energy_forecast %>% 
  autoplot()
```

```{r}
# or plot it added to the original data
energy_forecast %>% 
  autoplot(energy_ts)
```

## Assessing residuals

First, use broom::augment() to get the predicted values & residuals:

```{r}
# Append the predicted values (and residuals) to original energy data
energy_predicted <- broom::augment(energy_fit)

# use View(energy_predicted) to see the resulting df
```

```{r}

# make sure they look like pretty good predictions! (they do)
ggplot(data = energy_predicted) +
  geom_line(aes(x = date, y = res_total)) +
  geom_line(aes(x = date, y = .fitted), color = "red")
```

Now look at residuals

```{r}
ggplot(data = energy_predicted, aes(x = .resid)) +
  geom_histogram()
```

## See example html for other estimation methods of time series